---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MS2extract

<!-- badges: start -->
[![R-CMD-check](https://github.com/CooperstoneLab/MS2extract/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CooperstoneLab/MS2extract/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/CooperstoneLab/MS2extract/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/CooperstoneLab/MS2extract/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

The goal of MS2extract is to create a tool to import MS2 data of known standards
or material and targeted extract the MS2 expectra in order to create an
in-house MS2 library.

## Installation

You can install the development version of MS2extract from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("CooperstoneLab/MS2extract")
```

## Example

### PPM calculation

with ppm_range you can compute a the ppm error given a m/z value and a ppm
tolerance.

```{r example, eval=TRUE, include=TRUE}
library(MS2extract)
## calculating ppm range
chlorogenic_acid_pos <- 355.1023
ppm_error = 10
# Calculate ranges
ppm_range(mz = chlorogenic_acid_pos, ppm = ppm_error)
```

### Importing mzxml files

You can  import mzxml files by `import_mzxml()` function. This will import
your data into a data frame. In this case we are going to work with the
Procyanidin A2 spectra in neg mode with a collision energy of 20eV.

```{r}
ProcA2_file <- system.file("extdata",
                           "ProcyanidinA2_neg_20eV.mzXML",
                           package = "MS2extract")

ProcA2_raw <- import_mzxml(ProcA2_file)
```

### Extracting Procyanidin A2 spectra

In this case, the mzxml file contains only the scans of this metabolite and
the `extract_MS2()` functions exports the most intense MS2 scans, and displays
the TIC indicating which scan is using to export the MS2 data.

```{r}
extract_MS2(ProcA2_raw)
```




