---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MS2extract

<!-- badges: start -->
[![R-CMD-check](https://github.com/CooperstoneLab/MS2extract/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CooperstoneLab/MS2extract/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/CooperstoneLab/MS2extract/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/CooperstoneLab/MS2extract/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

The goal of MS2extract is to create a tool to import MS2 data of known standards
or material and targeted extract the MS2 expectra in order to create an
in-house MS2 library.

## Installation

You can install the development version of MS2extract from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("CooperstoneLab/MS2extract")
```

## Example

### PPM calculation

with ppm_range you can compute a the ppm error given a m/z value and a ppm
tolerance.

```{r example, eval=TRUE, include=TRUE}
library(MS2extract)
## calculating ppm range
chlorogenic_acid_pos <- 355.1023
ppm_error = 10
# Calculate ranges
ppm_range(mz = chlorogenic_acid_pos, ppm = ppm_error)
```

### Importing mzxml files

You can  import mzxml files by `import_mzxml()` function. This will import
your data into a data frame. In this case we are going to work with the
Procyanidin A2 spectra in neg mode with a collision energy of 20eV.

```{r}
# File path
ProcA2_file <- system.file("extdata",
                           "ProcyanidinA2_neg_20eV.mzXML",
                           package = "MS2extract")
# Compound data
ProcA2_data <- data.frame(Formula = "C30H24O12",Ionization_mode = "Negative")

ProcA2_raw <- import_mzxml(ProcA2_file, ProcA2_data)
```

### Extracting Procyanidin A2 spectra

In this case, the mzxml file contains only the scans of this metabolite and
the `extract_MS2()` functions exports the most intense MS2 scans, and displays
the TIC indicating which scan is using to export the MS2 data.

```{r message=FALSE, warning=FALSE}
ProcA2_ext <- extract_MS2(ProcA2_raw)
```

This function returns a list, which contains the MS2 spectra, and the TIC plot.
Since the extracted MS2 spectra contains low intensity signals (backgroun)
noise, you can remove them by using the `detect_mass()` function.

The raw MS2 spectra contains 6226 signals which the majority of them are low
intensity signals (background noise).

```{r}
dim(ProcA2_ext$MS2_spec)
```



### Detecting masses

The objective is to remove low intensity signals, you can detec masses by
allowing normalization by the base peak and then filtering the signals by 
a minimum treshold, or detecting mass with the abolute signal intensity.

Here, we will detect masses by normalization and with a minimum intensity of 
1%

```{r}
ProcA2_detected <- detect_mass(ProcA2_ext$MS2_spec,
                               normalize = TRUE, # Allow normalization
                               min_int = 1) # 1% as minimum intensity

dim(ProcA2_detected)
```

Now we can see that the x-axis in the MS2 spectra is reduced to 575 m/z since
the low intensity signals were removed.

```{r message=FALSE, warning=FALSE}
plot_MS2spectra(ProcA2_detected)
```


Although the resulting MS2 spectra contains 38 signals, for displaying
purposes we are going to include signlas greater than 10% intensity

```{r}
dplyr::filter(ProcA2_detected, intensity > 10)
```

### Comparing two MS2 spectra with cosine score

The purpose of this section is to compare two MS2 spectra that were extracted
using this package and using PCDL, a compound library developed by Agilent.

In this case, we include the MS2 spectra extracted with PCDL and use cosine
similarity score to compare them by the `compare_spectra()` function.

```{r}
# MS2 PCDL spectra
ProcA2_pcdl_fl <- system.file("extdata",
                              "ProcA2_neg_20eV_PCDL.csv",
                              package = "MS2extract")
# Reading the spectra
ProcA2_PCDL <- read.csv(ProcA2_pcdl_fl)

# Comparing both spectra
compare_spectra(ProcA2_detected, ProcA2_PCDL)

```


